---
title: "Project"
author: "Adam Del Rio and Nicholas Elich"
date: "2025-06-06"
output: pdf_document
---

### Packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(baseballr)
library(brms)
library(sn)
library(bayesplot)
library(tidybayes)
```

### Scraping Data

```{r}
start_date <- as.Date("2025-03-27")
end_date <- as.Date("2025-05-29")

# Initialize an empty list to store data frames
all_data <- list()

# Loop over the date range in 5-day increments
current_start <- start_date

while (current_start <= end_date) {
  current_end <- current_start + 4  # 5 days window
  
  # Make sure current_end doesn't exceed end_date
  if (current_end > end_date) {
    current_end <- end_date
  }
  
  # Fetch data for current 5-day window
  cat("Fetching data from", current_start, "to", current_end, "\n")
  temp_data <- statcast_search(start_date = as.character(current_start),
                               end_date = as.character(current_end),
                               player_type = "batter")
  
  # Append to list
  all_data <- append(all_data, list(temp_data))
  
  # Move to the next window
  current_start <- current_end + 1
}

# Combine all data frames into one
final_data <- bind_rows(all_data)
```

### NL East Player Names

```{r}
player_names <- c(
  "Murphy, Sean",
  "Olson, Matt",
  "Albies, Ozzie",
  "Allen, Nick",
  "Riley, Austin",
  "Verdugo, Alex",
  "Harris II, Michael",
  "White, Eli",
  "Ozuna, Marcell",
  "Baldwin, Drake",
  "Kelenic, Jarred",
  "De La Cruz, Bryan",
  "Fairchild, Stuart",
  "Arcia, Orlando",
  "Acuña Jr., Ronald",
  "Profar, Jurickson",
  "Tromp, Chadwick",
  "Williams, Luke",
  "Rosario, Eddie",
  "Torrens, Luis",
  "Alonso, Pete",
  "Acuña, Luisangel",
  "Lindor, Francisco",
  "Vientos, Mark",
  "Nimmo, Brandon",
  "Taylor, Tyrone",
  "Soto, Juan",
  "Marte, Starling",
  "Baty, Brett",
  "McNeil, Jeff",
  "Alvarez, Francisco",
  "Winker, Jesse",
  "Senger, Hayden",
  "Siri, Jose",
  "Azócar, José",
  "Young, Jared",
  "Ruiz, Keibert",
  "Lowe, Nathaniel",
  "García Jr., Luis",
  "Abrams, CJ",
  "Tena, José",
  "Wood, James",
  "Young, Jacob",
  "Crews, Dylan",
  "Bell, Josh",
  "Call, Alex",
  "Rosario, Amed",
  "DeJong, Paul",
  "Nuñez, Nasim",
  "Adams, Riley",
  "Hassell III, Robert",
  "Lile, Daylen",
  "Lipscomb, Trey",
  "Hicks, Liam",
  "Wagaman, Eric",
  "López, Otto",
  "Edwards, Xavier",
  "Norby, Connor",
  "Stowers, Kyle",
  "Myers, Dane",
  "Sánchez, Jesús",
  "Ramírez, Agustín",
  "Mervis, Matt",
  "Sanoja, Javier",
  "Pauley, Graham",
  "Hill, Derek",
  "Conine, Griffin",
  "Fortes, Nick",
  "Simón, Ronny",
  "Bride, Jonah",
  "Mesa, Víctor",
  "Brantly, Rob",
  "Realmuto, J.T.",
  "Harper, Bryce",
  "Stott, Bryson",
  "Turner, Trea",
  "Bohm, Alec",
  "Kepler, Max",
  "Rojas, Johan",
  "Castellanos, Nick",
  "Schwarber, Kyle",
  "Marsh, Brandon",
  "Sosa, Edmundo",
  "Marchán, Rafael",
  "Wilson, Weston",
  "Stevenson, Cal",
  "Clemens, Kody"
)


data2 <- final_data %>%
  filter(description == "hit_into_play" &
         (home_team %in% c("PHI", "ATL", "WSH", "NYM", "MIA") |
          away_team %in% c("PHI", "ATL", "WSH", "NYM", "MIA")) &
          player_name %in% player_names)
```

```{r}
hits <- data2 %>% group_by(player_name) %>%
  summarize(n = n()) %>%
  filter(n >= 40)

data3 <- data2 %>%
  inner_join(hits, by = "player_name")
```

### Reading in CSV file after initial scrape

```{r}
data3 <- read_csv("/Users/nicholas.elich/Downloads/NLEast2025Pitches.csv")
```

### Setting Priors and Tuning with Prior Predictive Distribution

```{r}
data4 <- data3 %>%
  mutate(across(c(release_speed, bat_speed, swing_length, launch_angle), scale))

x = 0.01
n_rep = 100000

beta0 = rnorm(n_rep, 90, 2.5)
sigma = rnorm(n_rep, 8, 3)
alpha = rnorm(n_rep, 0, 1.5)
beta_rel_speed = rnorm(n_rep, 1, 0)
beta_launch = rnorm(n_rep, 0, 0.6)
beta_swing_length = rnorm(n_rep, 6, 3)
beta_bat_speed = rnorm(n_rep, 3, 4)

y = rsn(n_rep, beta0 + beta_rel_speed*x + beta_launch*x + beta_swing_length*x + beta_bat_speed*x, sigma, alpha)

hist(y, xlab="Predicted Exit Velocity (mph)", main="Prior Predictive Tuning for Exit Velocity Based on Avg. Predictor Values")
summary(y)
```

### Checking Linear Regression Assumptions

```{r}
plot(lm(launch_speed ~ release_speed + bat_speed + swing_length + launch_angle,
  data = data4))
```

### Model with Our Own Priors

```{r}
model1 <- brm(
  launch_speed ~ release_speed + bat_speed + swing_length + launch_angle,
  data = data4,
  prior = c(prior(normal(90, 2.5), class = Intercept),
            prior(normal(8, 3), class = sigma),
            prior(normal(0, 1.5), class = alpha),
            prior(normal(1, 0.5), class="b", coef = "release_speed"),
            prior(normal(0, 0.6), class="b", coef="bat_speed"),
            prior(normal(6, 3), class="b", coef="swing_length"),
            prior(normal(3, 4), class="b", coef="launch_angle")),
  family = skew_normal(),
  chains = 4,
  iter = 3500,
  warmup = 1000, 
  seed = 1234
)

summary(model1)
get_prior(model1)
plot(model1)
```

```{r}
posterior_summary(model1, probs = c(0.10, 0.90))
```

### Model with brms Priors

```{r}
model2 <- brm(
  launch_speed ~ release_speed + bat_speed + swing_length + launch_angle,
  data = data4,
  family = skew_normal(),
  chains = 4,
  iter = 3500,
  warmup = 1000, 
  seed = 1234
)

summary(model2)
get_prior(model2)
plot(model2)
```

```{r}
posterior_summary(model2, probs = c(0.10, 0.90))
```

### Model with brms Priors and Hierarchical Component (Player Name)

```{r}
model3 <- brm(
  launch_speed ~ release_speed + bat_speed + swing_length + launch_angle + (1 | player_name),
  data = data4,
  family = skew_normal(),
  chains = 4,
  iter = 3500,
  warmup = 1000, 
  seed = 1234
)

summary(model3)
get_prior(model3)
plot(model3)
```

```{r}
posterior_summary(model3, probs = c(0.10, 0.90))
```

### Plotting the Random Effects of Players (Top 10 and Bottom 10)

```{r}
random_effects <- ranef(model3)$player_name[, , "Intercept"]

player_effects <- as.data.frame(random_effects) |> 
  rownames_to_column("player_name") |> 
  rename(
    estimate = Estimate,
    lower = Q2.5,
    upper = Q97.5
  ) |> 
  arrange(desc(estimate)) |> 
  slice_head(n=10)

ggplot(player_effects, aes(x = reorder(player_name, estimate), y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  coord_flip() +
  labs(
    x = "Player",
    y = "Estimated Random Intercept",
    title = "Player-Specific Effects from brms"
  ) +
  theme_minimal()

player_effects <- as.data.frame(random_effects) |> 
  rownames_to_column("player_name") |> 
  rename(
    estimate = Estimate,
    lower = Q2.5,
    upper = Q97.5
  ) |> 
  arrange(desc(estimate)) |> 
  slice_tail(n=10)

ggplot(player_effects, aes(x = reorder(player_name, estimate), y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2) +
  coord_flip() +
  labs(
    x = "Player",
    y = "Estimated Random Intercept",
    title = "Player-Specific Effects from brms"
  ) +
  theme_minimal()
```

### Posterior Predictive Checking for Model 1

```{r}
posterior_1 = model1 |> 
  spread_draws(b_Intercept, b_bat_speed, b_launch_angle, b_release_speed, b_swing_length, alpha, sigma) |> 
  mutate(mu_01 = b_Intercept + 0.01*(b_bat_speed + b_launch_angle + b_release_speed + b_swing_length))

posterior_1 <- posterior_1 |> 
  mutate(y_pred_01 = mu_01 + sigma * rnorm(nrow(posterior_1), 0, 1))

posterior_1 <- data.frame(posterior_1)

posterior_1 |> ggplot(aes(x=y_pred_01)) + 
  geom_histogram(aes(y=after_stat(density)), bins = 100, col="pink", fill="white") + 
  geom_density() + 
  theme_bw() + 
  ylab("Density") + 
  xlab("Predicted Exit Velocity for Average (0.01) Predictor Values")

quantile(posterior_1$y_pred_01, c(0.025, 0.1, 0.9, 0.975))

pp_check(model1, ndraw=100)
```

### Posterior Predictive Checking for Model 2

```{r}
posterior_2 = model2 |> 
  spread_draws(b_Intercept, b_bat_speed, b_launch_angle, b_release_speed, b_swing_length, alpha, sigma) |> 
  mutate(mu_01 = b_Intercept + 0.01*(b_bat_speed + b_launch_angle + b_release_speed + b_swing_length))

posterior_2 <- posterior_2 |> 
  mutate(y_pred_01 = mu_01 + sigma * rnorm(nrow(posterior_2), 0, 1))

posterior_2 <- data.frame(posterior_2)

posterior_2 |> ggplot(aes(x=y_pred_01)) + 
  geom_histogram(aes(y=after_stat(density)), bins = 100, col="pink", fill="white") + 
  geom_density() + 
  theme_bw() + 
  ylab("Density") + 
  xlab("Predicted Exit Velocity for Average (0.01) Predictor Values")

quantile(posterior_2$y_pred_01, c(0.025, 0.1, 0.9, 0.975))

pp_check(model2, ndraw=100)
```

### Posterior Predictive Checking for Model 3

```{r}
posterior_3 = model3 |> 
  spread_draws(b_Intercept, b_bat_speed, b_launch_angle, b_release_speed, b_swing_length, alpha, sigma) |> 
  mutate(mu_01 = b_Intercept + 0.01*(b_bat_speed + b_launch_angle + b_release_speed + b_swing_length))

posterior_3 <- posterior_3 |> 
  mutate(y_pred_01 = mu_01 + sigma * rnorm(nrow(posterior_3), 0, 1))

posterior_3 <- data.frame(posterior_3)

posterior_3 |> ggplot(aes(x=y_pred_01)) + 
  geom_histogram(aes(y=after_stat(density)), bins = 100, col="pink", fill="white") + 
  geom_density() + 
  theme_bw() + 
  ylab("Density") + 
  xlab("Predicted Exit Velocity for Average (0.01) Predictor Values")

quantile(posterior_3$y_pred_01, c(0.025, 0.1, 0.9, 0.975))

pp_check(model3, ndraw=100)
```
